---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

# Salivary gland tumor genetics project

Reported Telomere length and TERT promoter mutation in POR already. 
Preparation for writing new paper dealing with TL vs mtDNA, histopatholgy vs size


## Data load & process
```{r, echo=TRUE, message=FALSE, warning=FALSE}
Sys.setlocale("LC_CTYPE", "ko_KR.UTF-8") ##hangul data process
setwd("~/Documents/DataAnalysis/Salivary")
d <- read.csv("Saliva_TERT_pik3ca.csv")
# naming the histology group
d$hist_group<-factor(d$hist_group, levels=c(2,1,3), labels = c("WT", "PA", "CA")) # to exclude "etc" group in latter(heejin's opinion)
d <- subset(d, d$hist_group!="etc")
d$sex <- factor(d$sex, levels=c(1,2), labels = c("male", "female")) 
d$MT_copynumber <- d$MT_copynumber + 0.01 #for log scaling
d[15:16,16] = NA #exclude outlier
## Linear regression for oncogene vs size.

# Benign vs carcinoma subsetting for linear regression
d_PA <- subset(d, hist_group == "PA")
d_WT <- subset(d, hist_group == "WT")
d_CA <- subset(d, hist_group == "CA")
```

## Loading librarys
```{r, echo=TRUE, message=FALSE, warning=FALSE}
require("ggplot2")
require("moonBook")
require("ztable")
require("gridExtra")
```

## DATA summary
```{r, echo=TRUE, message=FALSE, warning=FALSE}
summary(d$sex)
summary(factor(d$location))
summary(d$hist_group)
19/32*100
8/32*100
5/32*100
mean(d$size) 
sd(d$size)
```

## Table 1
### Histology_grouping table
```{r, echo=TRUE, warning=FALSE}
d_table<-d[c("sex", "age", "location", "hist_group","size" )]
mytable(hist_group~., data=d_table)
```

## Figure 1(use vertical arrange)
### Box plot for ongene in salivary gland.
```{r, echo=TRUE, fig.height=8, fig.width=5, message=FALSE, warning=FALSE}
#TL(previously published at POR letter, however heejin insisted including)
fig1_tl <- ggplot(d, aes(hist_group, telomere_length)) + geom_boxplot(fill = "grey") +
  scale_x_discrete(labels = c ("WT","PA","CA","etc")) + 
  labs(x ="Histology group", y="Telomere length") + 
  theme(legend.position="none") +
  theme_bw() + scale_y_log10()

#PIK3CA
fig1_a <- ggplot(d, aes(hist_group, PIK3CA_amp)) + geom_boxplot(fill = "grey") +
  scale_x_discrete(labels = c ("WT","PA","CA","etc")) + 
  labs(x ="Histology group", y="PIK3CA amplication") + 
  theme(legend.position="none") +
  theme_bw()

#mitochondria CNV
fig1_b <- ggplot(d, aes(hist_group, MT_copynumber)) + geom_boxplot(fill="grey") +
  scale_x_discrete(labels = c ("WT","PA","CA","etc")) + 
  labs(x ="Histology group", y="mitochondria copy number") + 
  theme(legend.position="none") +
  theme_bw() + scale_y_log10()

figure1_vert <- arrangeGrob(fig1_tl, fig1_a, fig1_b, nrow = 3)
plot(figure1_vert)
ggsave("Figure_1_tl.tiff", figure1_vert, units = "in", width = 5, height = 12, dpi = 300)
```

## Figure 2
###plotting size vs oncogene, faceting plot
```{r, echo=TRUE, fig.height=6.5, fig.width=5, message=FALSE, warning=FALSE}
fig2_a <- ggplot(d, aes(size, telomere_length)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group) + 
  labs(x = NULL , y="Telomere length", shape = NULL) + 
  theme_bw() # + theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  # legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

fig2_b <- ggplot(d, aes(size, PIK3CA_amp)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  facet_wrap(~hist_group) +
  labs(x = NULL , y="PIK3CA amplification", shape = NULL) + 
  theme_bw()

fig2_c <- ggplot(d, aes(size, MT_copynumber)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  facet_wrap(~hist_group) +
  labs(x ="Size (cm)", y="Mitochondria copy number", shape = NULL) + 
  theme_bw() + scale_y_log10()

figure2 <- arrangeGrob(fig2_a, fig2_b, fig2_c, nrow = 3)
plot(figure2)
ggsave("Figure_2.tiff", figure2, units = "in", width = 5, height = 6.5, dpi = 300)
```

## Figure 2 correlation (Significant!!)
### size correlation spearman test in benign
```{r, echo=TRUE, message=TRUE, warning=FALSE}
cor.test(d_PA$size, d_PA$telomere_length, method = "spearman")
cor.test(d_WT$size, d_WT$telomere_length, method = "spearman")
cor.test(d_CA$size,d_CA$telomere_length, method="spearman")

cor.test(d_PA$size, d_PA$PIK3CA_amp, method = "spearman")
cor.test(d_WT$size, d_WT$PIK3CA_amp, method = "spearman")
cor.test(d_CA$size,d_CA$PIK3CA_amp, method="spearman")

cor.test(d_PA$size, d_PA$MT_copynumber, method = "spearman")
cor.test(d_WT$size, d_WT$MT_copynumber, method = "spearman")
cor.test(d_CA$size,d_CA$MT_copynumber, method="spearman")
```

## Figure 3
### plotting age vs telomere_length, oncogenes, facetting plot
```{r, echo=TRUE, fig.height=6.5, fig.width=5, message=FALSE, warning=FALSE}
fig3_a <- ggplot(d, aes(age, telomere_length)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group, scales = "free_x") + 
  labs(x ="age", y="Telomere length", shape = NULL) + 
  theme_bw() #+ theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  #legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

fig3_b <- ggplot(d, aes(age, PIK3CA_amp)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group, scales = "free_x") + 
  labs(x ="age", y="PIK3CA", shape = NULL) + 
  theme_bw() #+ theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  #legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

fig3_c <- ggplot(d, aes(age, MT_copynumber)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group, scales = "free_x") + 
  labs(x ="age", y="mitochondria", shape = NULL) + scale_y_log10() +
  theme_bw() # + theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  # legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

figure3 <- arrangeGrob(fig3_a, fig3_b, fig3_c, nrow = 3)
plot(figure3)
ggsave("Figure_3.tiff", figure3, units = "in", width = 5, height = 6.5, dpi = 300)
```

### age correlation spearman test in benign and malignant.
```{r, echo=TRUE, message=FALSE}
cor.test(d_PA$age, d_PA$telomere_length, method = "spearman")
cor.test(d_WT$age, d_WT$telomere_length, method = "spearman")
cor.test(d_CA$age, d_CA$telomere_length, method="spearman")

cor.test(d_PA$age, d_PA$PIK3CA_amp, method = "spearman")
cor.test(d_WT$age, d_WT$PIK3CA_amp, method = "spearman")
cor.test(d_CA$age, d_CA$PIK3CA_amp ,method="spearman")

cor.test(d_PA$age, d_PA$MT_copynumber, method = "spearman")
cor.test(d_WT$age, d_WT$MT_copynumber, method = "spearman")
cor.test(d_CA$age, d_CA$MT_copynumber ,method="spearman")
```

## Figure 4
### plotting telomere length vs MT_copynumber, histology group
```{r, echo=TRUE, message=FALSE}

ggplot(d, aes(telomere_length, MT_copynumber)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  stat_smooth(method = "lm", col = "black") +
  facet_wrap(~hist_group) + 
  labs(x ="Telomere length", y="Mitochondria copy number", shape = NULL) + 
  theme_bw() + scale_x_log10() + scale_y_log10() #+ 
  #theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  #legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

ggsave("Figure_4.tiff", units = "in", width = 6, height = 3, dpi =300)
```

### spearman test
```{r, echo=T, message=F}
cor.test(d_PA$MT_copynumber, d_PA$telomere_length, method = "spearman")
cor.test(d_WT$MT_copynumber, d_WT$telomere_length, method = "spearman")
cor.test(d_CA$MT_copynumber, d_CA$telomere_length, method = "spearman")
```