---
output: html_document
runtime: shiny
---
# Salivary gland tumor genetics project

### Git control
Open the shell.  
git push -u origin master  
and login the git.  

### Data load
```{r echo=FALSE}
Saliva <- read.csv("~/Documents/R/Salivary/Saliva_TERT_pik3ca.csv")
```


```{r echo=FALSE}
# loading librarys
require("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
require("moonBook", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
require("ztable", lib.loc="/Library/Frameworks/R.framework/Versions/3.2/Resources/library")
```


### Data summary
```{r echo=FALSE}
summary(Saliva)
```


```{r echo=FALSE}
#data manipulation 
Saliva$hist_group<-factor(Saliva$hist_group, levels=c(1,2,3), labels = c("pleomorphic adenoma", "warthin", "carcinoma")) # naming the histology group
Saliva$sex <- factor(Saliva$sex, levels=c(1,2), labels = c("male", "female")) # naming the gender
```


```{r echo=FALSE}
sexvar<-c("sex", "age", "location", "hist_group","size" )
#Saliva_table<-Saliva[sexvar] #create new data set for creating mytable
#mytable(sex~.,data=Saliva_table) #moon's package
```


```{r echo=FALSE}
#res_hist= mytable(hist_group~., data=Saliva_table)
#ztable(res_hist)
#mytable(hist_group~., data=Saliva_table) 
```

### Gender table
```{r echo=FALSE}
Saliva_3 <- subset(Saliva, hist_group != "etc") # creating data subset excluing etc group of histolgy.
Saliva_3_table <- Saliva_3[sexvar]
#res_sex_3= mytable(sex~.,data=Saliva_3_table)
#ztable(res_sex_3)
#res_hist_3= mytable(hist_group~., data=Saliva_3_table)
#ztable(res_hist_3)
mytable(sex~.,data=Saliva_3_table)
```

### Histology-based table
```{r echo=FALSE}
mytable(hist_group~., data=Saliva_3_table)
```

### Box plot
```{r echo=FALSE}
Saliva_3_TL <- subset(Saliva_3, telomere_length<40) # excluding outlier data in the telomere length
qplot(hist_group, size, data = Saliva_3, geom = c("boxplot", "jitter"), fill = hist_group) # plotting the box plot for size
qplot(hist_group, telomere_length, data=Saliva_3_TL, geom=c("boxplot", "jitter"), fill=hist_group) # plotting the box plot for telomere length
qplot(hist_group, PIK3CA_amp, data=Saliva_3, geom=c("boxplot", "jitter"), fill=hist_group) # plotting the box plot for PIK3CA expression 
Saliva_3_MT <- subset(Saliva_3, MT_copynumber<100) # excluding outlier data in the mitochondria copy number
qplot(hist_group, MT_copynumber, data=Saliva_3_MT, geom=c("boxplot", "jitter"), fill=hist_group) # plotting the box plot for mitochondria copy number
```

### ANOVA
Size vs histology
```{r echo=FALSE}
Saliva.size.aov <- aov(size~hist_group, data=Saliva_3)
summary(Saliva.size.aov)
```

Telomere length vs histology
```{r echo=FALSE}
Saliva.tl.aov <- aov(telomere_length~hist_group, data=Saliva_3_TL)
summary(Saliva.tl.aov)
```

PIK3CA expression vs histology
```{r echo=FALSE}
Saliva.PIK.aov <- aov(PIK3CA_amp~hist_group, data=Saliva_3)
summary(Saliva.PIK.aov)
```

mitochondria vs histology
```{r echo=FALSE}
Saliva.MT.aov <- aov(MT_copynumber~hist_group, data=Saliva_3_MT)
summary(Saliva.MT.aov)
```

### ploting size vs telomere lengh , histology group
```{r echo=FALSE}
qplot(size, telomere_length, data=Saliva_3, shape=hist_group, ylim = c(0,40))
```

### ploting size vs PIK3CA expression , histology group
```{r echo=FALSE}
qplot(size, PIK3CA_amp, data=Saliva_3, shape=hist_group)
```

### ploting size vs Mitochondria copy number, histology group
```{r echo=FALSE}
qplot(size, MT_copynumber, data=Saliva_3, shape = hist_group, ylim = c(0,9))
```
