---
output: html_document
runtime: shiny
---
# Salivary gland tumor genetics project

I am writing a paper descring these result at google docs.


### Data load
```{r echo=FALSE}
d <- read.csv("Saliva_TERT_pik3ca.csv")
```


```{r echo=FALSE}
# loading librarys
require("ggplot2")
require("moonBook")
require("ztable")
```

### Data summary and manupulation
```{r echo=FALSE}
summary(d)

# naming the histology group
d$hist_group<-factor(d$hist_group, levels=c(1,2,3,4), labels = c("pleomorphic adenoma", "warthin", "carcinoma", "etc")) 
d$sex <- factor(d$sex, levels=c(1,2), labels = c("male", "female")) 
```


### Gender table
```{r echo=FALSE}
sexvar<-c("sex", "age", "location", "hist_group","size" )
d_t<-d[sexvar]

ztable( mytable(sex~.,data=d_t))
```

### Histology table
```{r echo=FALSE}
ztable(mytable(hist_group~., data=d_t))
```


### Box plot
```{r echo=FALSE}
qplot(hist_group, size, data = d, geom = c("boxplot"), fill = hist_group) # plotting the box plot for size
qplot(hist_group, telomere_length, data=d, geom=c("boxplot"), fill=hist_group) # plotting the box plot for telomere length
qplot(hist_group, PIK3CA_amp, data=d, geom=c("boxplot"), fill=hist_group) # plotting the box plot for PIK3CA expression
d_mt <-subset(d, MT_copynumber <100)

qplot(hist_group, MT_copynumber, data=d_mt, geom=c("boxplot"), fill=hist_group) # plotting the box plot for mitochondria copy number
```

### ANOVA
Size vs histology
```{r echo=FALSE}
d_size_aov <- aov(size~hist_group, data=d)
summary(d_size_aov)
```

Telomere length vs histology
```{r echo=FALSE}
d_tl_aov <- aov(telomere_length~hist_group, data=d)
summary(d_tl_aov)
```

PIK3CA expression vs histology
```{r echo=FALSE}
d_PIK_aov <- aov(PIK3CA_amp~hist_group, data=d)
summary(d_PIK_aov)
```

mitochondria vs histology
```{r echo=FALSE}
d_mt_aov <- aov(MT_copynumber~hist_group, data=d_mt)
summary(d_mt_aov)
```

### ploting size vs telomere length , histology group
```{r echo=FALSE}
qplot(size, telomere_length, data=d, shape=hist_group, color = hist_group, ylim = c(0,120))
```

### ploting size vs PIK3CA expression , histology group
```{r echo=FALSE}
qplot(size, PIK3CA_amp, data=d, shape=hist_group, color = hist_group)
```

### ploting size vs Mitochondria copy number, histology group
```{r echo=FALSE}
qplot(size, MT_copynumber, data=d, shape = hist_group, color = hist_group, ylim = c(0,10))
```

### ploting telomere length vs PIK3CA, histology group
```{r echo=FALSE}
qplot(telomere_length, PIK3CA_amp, data=d,  shape=hist_group, color = hist_group, xlim = c(0,120))
```

### ploting telomere length vs mitochondria, histology group
```{r echo=FALSE}
qplot(telomere_length, MT_copynumber, data=d,  shape=hist_group, color = hist_group, xlim = c(0,120), ylim = c(0,9))
```
