---
output: html_document
runtime: html
---
# Salivary gland tumor genetics project

Reported Telomere length and TERT promoter mutation in POR already. 
Preparation for writing new paper dealing with TL vs mtDNA, histopatholgy vs size

=======
### Data load
```{r, include=FALSE, message=FALSE}
Sys.setlocale("LC_CTYPE", "ko_KR.UTF-8") ##hangul data process
d <- read.csv("Saliva_TERT_pik3ca.csv")
# naming the histology group
d$hist_group<-factor(d$hist_group, levels=c(1,2,3), labels = c("PA", "WT", "CA")) # to exclude "etc" group in latter(heejin's opinion)
d$sex <- factor(d$sex, levels=c(1,2), labels = c("male", "female")) 
d$MT_copynumber <- d$MT_copynumber+ 0.01
d[15:16,16] = NA #exclude outlier
# data tagging for facetting Benign vs CA
BenignCA <- factor(d$hist_group =="CA", levels=c(F,T), labels = c("Benign", "CA"))
d<-cbind(d, BenignCA)

```

```{r, echo=TRUE, message=FALSE}
# loading librarys
require("ggplot2")
require("moonBook")
require("ztable")
require("gridExtra")
# require("Rcmdr")
```

## data subsetting according to heejin's opinion
```{r, echo=TRUE, message=FALSE}
d <- subset(d, d$hist_group!="etc")
```



### Data summary and manupulation
```{r, echo=TRUE, message=FALSE}
summary(d)
sd(d$size)
```
Male 19, female 17. 
Parotid gland 30, Submandibular gland 6.
PA 19, WT 8, CA 5, etc 4.
size mean 3 (+-sd: 1.13)cm
benign 31/36(86.1%), malig 5/36(13.9%)

after exclude etc(benign)
Male 16, female 16
Parotid 27, Submandibular 5
PA 19, WT 8, CA 5
size mean 3.05(+-sd: 1.18)cm
benign 27/32(84.4%), CA 5/32(15.6%)

### Gender table
```{r, echo=TRUE, message=FALSE}
d_table<-d[c("sex", "age", "location", "hist_group","size" )]
# print(ztable(mytable(hist_group~.,data=d_table)))
# print(ztable(mytable(sex~.,data=d_table))) #excluded by heejin
```
### gender age compare with MW test
```{r, echo=TRUE, message=FALSE}
male_age <- subset(d$age, d$sex =="male")
female_age <- subset(d$age, d$sex == "female")
wilcox.test(male_age, female_age)
```
### gender size compare with MW test
```{r, echo=TRUE, message=FALSE}
male_size <- subset(d$size, d$sex =="male")
female_size <- subset(d$size, d$sex == "female")
wilcox.test(male_size, female_size)
```

### location table
```{r, echo=TRUE, message=FALSE}
# print(ztable(mytable(location~age + hist_group + size, data = d_table))) #excluded by heejin
```
### location age compare with MW test
```{r, echo=TRUE, message=FALSE}
paro_age <- subset(d$age, d$location_group == 1)
subman_age <- subset(d$age, d$location_group == 2)
wilcox.test(paro_age, subman_age)
```
### location size compare with MW test
```{r, echo=TRUE, message=FALSE}
paro_size <- subset(d$size, d$location_group == 1)
subman_size <- subset(d$size, d$location_group == 2)
wilcox.test(paro_size, subman_size)
```

## Table 1
### Histology_grouping table
```{r, echo=TRUE, message=FALSE}
mytable(hist_group~., data=d_table)
# print(ztable(mytable(hist_group~., data=d_table))) #table_1 for final version
# summary(aov(size~hist_group, data=d)) same result with chi-squire table analysis.
```

### histology all table.
```{r, echo=TRUE, message=FALSE}
# print(ztable(mytable(sex~histology, data=d_CA))) #excluded by heejin
# summary(aov(size~hist_group, data=d)) same result with chi-squire table analysis.
```
### Oncogene table
```{r, echo=TRUE, message=FALSE}
d_onco_t<-d[c("hist_group", "MT_copynumber", "PIK3CA_amp")] # telomere_length ommited due to POR letter
# print(ztable(mytable(hist_group~., data=d_onco_t))) #figur들로대체.

#ANOVA has same result chi-squire table analysis
#summary(aov(telomere_length~hist_group, data=d))
#summary(aov(PIK3CA_amp~hist_group, data=d))
#summary(aov(MT_copynumber~hist_group, data=d))
```

## Figure 1(use vertical arrange)
### Box plot for ongene in salivary gland.
```{r, echo=TRUE, message=FALSE}
# plotting the box plot for telomere length. POR letter published
# omit in the new paper.
#ggplot(d, aes(hist_group, telomere_length)) + geom_boxplot(fill = "grey") +
#  scale_x_discrete(labels = c ("PA","WT","CA","etc")) + 
#  labs(x ="Histology group", y="Telomere length") + 
#  theme(legend.position="none") +
#  theme_bw() + scale_y_log10()
#ggsave("telomere.tiff", units = "in", width = 5, height = 4, dpi =300) 


#PIK3CA
fig1_a <- ggplot(d, aes(hist_group, PIK3CA_amp)) + geom_boxplot(fill = "grey") +
  scale_x_discrete(labels = c ("PA","WT","CA","etc")) + 
  labs(x ="Histology group", y="PIK3CA amplication") + 
  theme(legend.position="none") +
  theme_bw()

# ggsave("Fig1_a.tiff", units = "in", width = 5, height = 4, dpi = 300)

#mitochondria CNV
fig1_b <- ggplot(d, aes(hist_group, MT_copynumber)) + geom_boxplot(fill="grey") +
  scale_x_discrete(labels = c ("PA","WT","CA","etc")) + 
  labs(x ="Histology group", y="mitochondria copy number") + 
  theme(legend.position="none") +
  theme_bw() + scale_y_log10()

figure1_vert <- arrangeGrob(fig1_a, fig1_b, nrow = 2)
figure1_hori <- arrangeGrob(fig1_a, fig1_b, ncol = 2)

ggsave("Figure_1.tiff", figure1_vert, units = "in", width = 5, height = 8, dpi = 300)
ggsave("Fig1_h.tiff", figure1_hori, units = "in", width = 10, height = 4, dpi = 300)
```

## Figure 2
###plotting size vs oncogene, faceting plot
```{r, echo=TRUE, message=FALSE}
fig2_a <- ggplot(d, aes(size, telomere_length)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group) + 
  labs(x = NULL , y="Telomere length", shape = NULL) + 
  theme_bw() # + theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  # legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

fig2_b <- ggplot(d, aes(size, PIK3CA_amp)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  facet_wrap(~hist_group) +
  labs(x = NULL , y="PIK3CA amplification", shape = NULL) + 
  theme_bw()

fig2_c <- ggplot(d, aes(size, MT_copynumber)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  facet_wrap(~hist_group) +
  labs(x ="Size (cm)", y="Mitochondria copy number", shape = NULL) + 
  theme_bw() + scale_y_log10()

figure2 <- arrangeGrob(fig2_a, fig2_b, fig2_c, nrow = 3)

ggsave("Figure_2.tiff", figure2, units = "in", width = 5, height = 6.5, dpi = 300)
```


## Linear regression for oncogene vs size.
```{r, echo=TRUE, message=FALSE}
# Benign vs carcinoma subsetting for linear regression
d_benign <- subset(d, hist_group != "CA")
d_PA <- subset(d, hist_group == "PA")
d_WT <- subset(d, hist_group == "WT")
d_CA <- subset(d, hist_group == "CA")
```
onc_size_be.lm <- lm(size ~ telomere_length + PIK3CA_amp + MT_copynumber, data = d_benign)
onc_size_PA.lm <- lm(size ~ telomere_length + PIK3CA_amp + MT_copynumber, data = d_PA)
onc_size_WT.lm <- lm(size ~ telomere_length + PIK3CA_amp + MT_copynumber, data = d_WT)
onc_size_CA.lm <- lm(size ~ telomere_length + PIK3CA_amp + MT_copynumber, data = d_CA)
summary(onc_size_be.lm)
summary(onc_size_PA.lm)
summary(onc_size_WT.lm)
summary(onc_size_CA.lm)

non-significant.
benign 
tl vs size p = 0.0522
pik3 vs size p = 0.8212
MT_CN vs size p = 0.6074
PA
tl vs size p = 0.1257
pik3 vs size p = 0.471828
MT_CN vs size p = 0.854790
WT
tl vs size p = 0.6962
pik3 vs size p = 0.1960
MT_CN vs size p = 0.1784
CA not available due to low number.


## Figure 2 correlation (Significant!!)
### size correlation spearman test in benign
```{r, echo=TRUE, message=FALSE}
# cor.test(d_benign$size,d_benign$telomere_length ,method="spearman")
cor.test(d_PA$size, d_PA$telomere_length, method = "spearman")
cor.test(d_WT$size, d_WT$telomere_length, method = "spearman")
cor.test(d_CA$size,d_CA$telomere_length, method="spearman")

# cor.test(d_benign$size,d_benign$PIK3CA_amp ,method="spearman")
cor.test(d_PA$size, d_PA$PIK3CA_amp, method = "spearman")
cor.test(d_WT$size, d_WT$PIK3CA_amp, method = "spearman")
cor.test(d_CA$size,d_CA$PIK3CA_amp ,method="spearman")

# cor.test(d_benign$size,d_benign$MT_copynumber ,method="spearman")
cor.test(d_PA$size, d_PA$MT_copynumber, method = "spearman")
cor.test(d_WT$size, d_WT$MT_copynumber, method = "spearman")
cor.test(d_CA$size,d_CA$MT_copynumber ,method="spearman")
```
non significant 
TL benign 0.08754 PA 0.03968 WT 0.9389 CA 0.95
pik3ca benign 0.5761 PA 0.6105 WT 0.01954 CA 0.417
mtDNA benign 0.759 PA 0.1805 WT 0.002439 CA 1

## Figure 3
### plotting age vs telomere_length, oncogenes, facetting plot
```{r, echo=TRUE, message=FALSE}
fig3_a <- ggplot(d, aes(age, telomere_length)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group, scales = "free_x") + 
  labs(x ="age", y="Telomere length", shape = NULL) + 
  theme_bw() #+ theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  #legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

fig3_b <- ggplot(d, aes(age, PIK3CA_amp)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group, scales = "free_x") + 
  labs(x ="age", y="PIK3CA", shape = NULL) + 
  theme_bw() #+ theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  #legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

fig3_c <- ggplot(d, aes(age, MT_copynumber)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) + 
  facet_wrap(~hist_group, scales = "free_x") + 
  labs(x ="age", y="mitochondria", shape = NULL) + scale_y_log10() +
  theme_bw() # + theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  # legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure


figure3 <- arrangeGrob(fig3_a, fig3_b, fig3_c, nrow = 3)

ggsave("Figure_3.tiff", figure3, units = "in", width = 5, height = 6.5, dpi = 300)
```

## Linear regression for oncogene vs age.
{r, echo=TRUE, message=FALSE}
onc_age_be.lm <- lm(age ~ telomere_length + PIK3CA_amp + MT_copynumber, data = d_benign)
summary(onc_age_be.lm)
onc_age_CA.lm <- lm(age ~ telomere_length + PIK3CA_amp + MT_copynumber, data = d_CA)
summary(onc_age_CA.lm)

non-significant
benign
tl vs age p = 0.259
CA not available due to low number. 

### age correlation spearman test in benign and malignant.
```{r, echo=TRUE, message=FALSE}
# cor.test(d_benign$size,d_benign$telomere_length ,method="spearman")
cor.test(d_PA$age, d_PA$telomere_length, method = "spearman")
cor.test(d_WT$age, d_WT$telomere_length, method = "spearman")
cor.test(d_CA$age, d_CA$telomere_length, method="spearman")

# cor.test(d_benign$size,d_benign$PIK3CA_amp ,method="spearman")
cor.test(d_PA$age, d_PA$PIK3CA_amp, method = "spearman")
cor.test(d_WT$age, d_WT$PIK3CA_amp, method = "spearman")
cor.test(d_CA$age, d_CA$PIK3CA_amp ,method="spearman")

# cor.test(d_benign$size,d_benign$MT_copynumber ,method="spearman")
cor.test(d_PA$age, d_PA$MT_copynumber, method = "spearman")
cor.test(d_WT$age, d_WT$MT_copynumber, method = "spearman")
cor.test(d_CA$age, d_CA$MT_copynumber ,method="spearman")
```
non-significant
TL PA 0.5308 WT 0.9084 CA 0.35
PIK3 PA 0.9935 WT 0.5284 CA 0.75
MT PA 0.3655 WT 0.6703 CA 0.6833

### plotting telomere length vs MT_copynumber, histology group
```{r, echo=TRUE, message=FALSE}

ggplot(d, aes(telomere_length, MT_copynumber)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  stat_smooth(method = "lm", col = "black") +
  facet_wrap(~hist_group) + 
  labs(x ="Telomere length", y="Mitochondria copy number", shape = NULL) + 
  theme_bw() + scale_x_log10() + scale_y_log10() #+ 
  #theme(legend.position = c(.99, .99), legend.justification = c("right", "top"),
  #legend.box.just = "right",legend.margin = margin(1, 1, 1, 1)) # draw legend inside the figure

ggsave("Figure_4.tiff", units = "in", width = 6, height = 3, dpi =300)
```

### Linear regression for telomere vs mt_copynumber
{r, echo=TRUE, message=FALSE}
tl_mt_be.lm <- lm(telomere_length ~ MT_copynumber, data = d_benign)
summary(tl_mt_be.lm)
tl_mt_PA.lm <- lm(telomere_length ~ MT_copynumber, data = d_PA)
summary(tl_mt_PA.lm)
tl_mt_CA.lm <- lm(telomere_length ~ MT_copynumber, data = d_CA)
summary(tl_mt_CA.lm)

non-significant.
benign p = 0.4345
PA p = 0.3
CA p = 0.838

## Figure 4 spearman(non-significant)
### tl vs mt correlation spearman test in benign and malignant.
```{r, echo=TRUE, message=FALSE}
cor.test(d_benign$telomere_length,d_benign$MT_copynumber ,method="spearman")
cor.test(d_PA$telomere_length, d_PA$MT_copynumber, method = "spearman")
cor.test(d_WT$telomere_length, d_WT$MT_copynumber, method = "spearman")
cor.test(d_CA$telomere_length,d_CA$MT_copynumber ,method="spearman")
```
non-significant.
benign p = 0.4551
PA p = 0.399
WT p = 0.5948
CA p = 0.95

#below here, analyzed but not including for paper
### ploting telomere length vs PIK3CA, histology group
```{r, echo=TRUE, message=FALSE}
ggplot(d, aes(telomere_length, PIK3CA_amp)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  stat_smooth(method = "lm", col = "black") +
  facet_wrap(~BenignCA) + 
  labs(x ="Telomere length", y="PIK3CA amplification", shape = NULL) + 
  theme_bw() + scale_x_log10()

ggsave("TLvsPI3.tiff", units = "in", width = 5, height = 4, dpi =300)
```


### plotting pik3ca vs mt_copynumber
```{r, echo=TRUE, message=FALSE}
ggplot(d, aes(PIK3CA_amp, MT_copynumber)) + geom_point(aes(shape = factor(hist_group)), show.legend = F) +
  stat_smooth(method = "lm", col = "black") +
  facet_wrap(~BenignCA) + 
  labs(x ="PIK3CA amplification", y="Mitochondria copy number", shape = NULL) + 
  theme_bw() + scale_y_log10()

ggsave("PI3vsMT.tiff", units = "in", width = 5, height = 4, dpi =300)
```

## linear regression for other oncogene.
```{r, echo=TRUE, message=FALSE}
tl_pik_be.lm <- lm(telomere_length ~ PIK3CA_amp, data = d_benign)
summary(tl_pik_be.lm)
tl_pik_CA.lm <- lm(telomere_length ~ PIK3CA_amp, data = d_CA)
summary(tl_pik_CA.lm)

pik_mt_be.lm <- lm(PIK3CA_amp ~ MT_copynumber, data = d_benign)
summary(pik_mt_be.lm)
pik_mt_CA.lm <- lm(PIK3CA_amp ~ MT_copynumber, data = d_CA)
summary(pik_mt_CA.lm)

```
non-significant.

### 3D plotting telomere length vs MT vs PIK3CA, histology group
{r echo=FALSE}
'
d_3d <- subset(d, telomere_length <120 & MT_copynumber <9)
z <- d_3d$telomere_length
y <- d_3d$MT_copynumber
x <- d_3d$PIK3CA_amp
scatter3d(x, y, z, surface = FALSE, 
          groups = d_3d$hist_group, 
          surface.col = c("red","green", "blue", "purple"),
          xlab = "PIK3CA ampication", ylab = "MT_copy number", zlab = "telomere length",
          axis.col = c("black", "black", "black"))
# require("rgl")          
# rgl.snapshot(filename = "TL_PI3_MT_3d.png")   그림을 내보기전에 축을적당히 회전시켜야함.
'



### drawing box plot sex vs age. omit.
{r, echo=TRUE, message=FALSE}
'
ggplot(d, aes(sex, age)) + geom_boxplot(aes(fill=sex)) +
  theme_bw() + scale_fill_brewer(palette = "Greys") + 
  theme(legend.position="none")
  #+  scale_fill_brewer(palette = "Set1")

ggsave("sex_age.tiff", units = "in", width = 5, height = 4, dpi =300)
'


### benign tumor histology, gender density bar graph. omit.
{r, echo=TRUE, message=FALSE}
'
d_b <- subset(d, hist_group != "carcinoma")

d_b_sub <-d_b[c("sex", "age", "location", "histology", "size")]

d_a_sub <- d[c("sex", "age", "location", "histology", "size")]

reorder_size <- function(x) {
  factor(x, levels = names(sort(table(x), decreasing = T)))
}

ggplot(d_a_sub, aes(reorder_size(histology))) + geom_bar(aes(fill=sex)) +
  scale_x_discrete("histology", 
                   labels = c("pleomorphic adenoma" ="PA", 
                              "warthin tumor" = "Warthin", 
                              "basal cell adenoma" = "BCA",
                              "epidermal inclusion cyst" = "EIC",
                              "chronic sclerosing sialadenitis" = "CSS",
                              "unremarkable" = "N/D")) + 
  scale_fill_brewer(palette = "Set1") +
  labs(x ="histolgy", y = "")

ggsave("histology_sex.tiff", units = "in", width = 5, height = 4, dpi =300)

ggplot(d_a_sub, aes(sex)) + geom_bar(aes(fill=reorder_size(histology))) +
  scale_fill_brewer(palette = "Spectral")

ggsave("sex_hist.tiff", units = "in", width = 5, height = 4, dpi =300)

ggplot(d_a_sub, aes(sex)) + geom_bar(aes(fill=location)) +
  scale_fill_brewer(palette = "Set1")

ggsave("sex_location.tiff", units = "in", width = 5, height = 4, dpi =300)

ggplot(d_a_sub, aes(x = sex,y =size)) + 
  geom_dotplot(binaxis = "y",binwidth = 0.2, stackdir = "center", aes(fill = sex)) +
  scale_fill_brewer(palette = "Set1")

ggsave("sex_size.tiff", units = "in", width = 5, height = 4, dpi =300)
'





### Histology vs size dot plot. omit.
{r echo=F}
'''
ggplot(d, aes(x = hist_group,y =size)) + 
  geom_dotplot(binaxis = "y",binwidth = 0.2, stackdir = "center", aes(shape = hist_group)) +
  scale_fill_brewer(palette = "Set1")

ggsave("histVSsize.tiff", units = "in", width = 5, height = 4, dpi =300)
'''


